---
title: "Group 3: work"
output: html_notebook
---

# Initialisation

```{r}
n <- 18                                 #size vector
p <- 20                                 #dimension vector
mu <- rep(0, p)                         #mean vector equal 0
Gamma <- 'diag<-'(matrix(0, p, p), 1)   #Gamma matrix p x p: Identity for testing

```

# random error epsilon_i in p dim 

Random vector uniformly distributed on the unit hypersphere in R^p.
[Muller , Marsaglia](http://extremelearning.com.au/how-to-generate-uniformly-random-points-on-n-spheres-and-n-balls/) ('Normalised Gaussians') method is used here.
```{r}
Ui <- function(x){
  #set.seed(0)          #Set seed for reproducibility
  v <- rnorm(p, 0, 1)
  norm <- norm(v, type = '2')
  U_i <- v / norm
  return(U_i)
}
```

Non negative random variable independent of Ui:  
```{r}
Ri <- function(x){
  #set.seed(0)          #Set seed for reproducibility
  R_i <- runif(1, 0, 100) # 100 is arbitrary
  return(R_i)
}
```

Special case Ri: Ri^2 ~ [Chi-square distribution](https://statisticsglobe.com/chi-square-distribution-in-r-dchisq-pchisq-qchisq-rchisq), p degree of freedom:
```{r}
Ri_Chi <- function(x){
  #set.seed(0)          #Set seed for reproducibility
  R_i <- rchisq(1, df = p) ** 0.5
  return(R_i)
}
```

Finally 
```{r}
epsilon_i <- function(x){
  epsi <- Gamma %*% (Ri_Chi() * Ui()) #nb: %*% is for matrix product
  #epsi <- Gamma %*% (Ri() * Ui())
  return(epsi)
}
```

# Generate data

First, create a file:
```{r}
file.create("C:/Users/maria/OneDrive/Documents/Unif 2021/mult. high dim. stat/DataTest.txt")
FilePath <- "C:/Users/maria/OneDrive/Documents/Unif 2021/mult. high dim. stat/DataTest.txt"    #To modify

```

This loop fill the file with the data. Each row is a vector of p dimension.
```{r}
for (i in 1:n){
  Xi <- mu + epsilon_i()
  write.table(t(Xi) , FilePath, append = TRUE, sep = " ", dec = ".",
              row.names = FALSE, col.names = FALSE) #nb: t() is for transpose
}
```

# Mean vector test Tn

This command import data from file and convert the data into a matrix:
```{r}
my_data <- matrix(scan(file = file.choose()),ncol=p,byrow=TRUE)

#my_data <- matrix(scan(file = FilePath),ncol=p,byrow=TRUE) #import data from file

```

This command create an empty matrix to fill with the vectors Zi:
```{r}
Z <- matrix(, nrow = n, ncol = p)
```

Each empty row of the previous matrix is replace by a Zi vector:
```{r}
for (i in 1:n){
  #matrix with the vector Zi. Each row is a vector of p components
  Z[i,] <- my_data[i,] / norm(my_data[i,], type= '2')
}
```

Finally, the Test statistic Tn:
```{r}
Test_n <- function(x){
  Tn <- 0               
  for (i in 2:n){
    for (j in 1:(i-1)){
      Tn <- Tn + Z[i,] %*% Z[j,]
  return(Tn)
    }
  }
}
```

